---
title: "Final Assignment"
author: "Tirani"
date: "December 3, 2021"
output:
  pdf_document: default
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)
library(car)
library(tidyverse)
library(ggplot2)
library(psych)
library(flextable)
library(ggpubr)
library(rstatix)
library(PMCMRplus)
library(xtable)
library(faraway)
library(phia)
library(effects)
library(MASS)
library(carData)
library(dplyr)
library(latex2exp)
library(SASxport)
library(nhanesA)
library(ggpmisc)
library(RColorBrewer)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
set_flextable_defaults(fonts_ignore = TRUE)

demographic_nhanes_1718 <- 
  read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/DEMO_J.xpt")
bodymeasure_nhanes_1718 <- 
  read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/BMX_J.xpt")
bp_chol_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/BPX_J.xpt")
diabetes_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/DIQ_J.xpt")
cardiovas_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/CDQ_J.xpt")
curr_health_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/HSQ_J.xpt")
income_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/INQ_J.xpt")
housing_char_nhanes_1718 <- 
  read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/HOQ_J.xpt")
food_security_nhanes_1718 <- 
  read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/FSQ_J.xpt")
diet_nutr_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/DBQ_J.xpt")

physical_act_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/PAQ_J.xpt")

smoke_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/SMQ_J.xpt")

sleep_disord_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/SLQ_J.xpt")

hsCRP_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/HSCRP_J.xpt")

HDL_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/HDL_J.xpt")
LDL_trigly_nhanes_1718 <- read.xport("~/Documents/R/stat_suni_1/R_1_suni/final_assignment/TRIGLY_J.xpt")

```

```{r}
View(demographic_nhanes_1718)
View(bodymeasure_nhanes_1718)
View(bp_chol_nhanes_1718)
View(diabetes_nhanes_1718)
View(cardiovas_nhanes_1718)
View(curr_health_1718)
View(income_nhanes_1718)
View(housing_char_nhanes_1718)
View(food_security_nhanes_1718)
View(diet_nutr_nhanes_1718)

View(physical_act_nhanes_1718)
View(smoke_nhanes_1718)
View(sleep_disord_nhanes_1718)
```

```{r Merge data}
#Merge dataframes
df_list <- list(demographic_nhanes_1718, bodymeasure_nhanes_1718, sleep_disord_nhanes_1718, hsCRP_nhanes_1718, HDL_nhanes_1718)

my_merge <- function(demographic_nhanes_1718, bodymeasure_nhanes_1718){                         merge(demographic_nhanes_1718, bodymeasure_nhanes_1718, by = "SEQN")
}

data_merge <- Reduce(my_merge, df_list)


nhanes_subset <- subset(data_merge, select = c(SEQN, RIAGENDR, RIDAGEYR, BMXWT, BMXHT, 
                                     BMXBMI, SLD012, LBXHSCRP, LBDHDD, RIDRETH3, 
                                     DMDEDUC2))

#Convert the values 7 (Refused) and 9 (Don’t Know) in education level (DMDEDUC2) to NA
nhanes_subset$DMDEDUC2[nhanes_subset$DMDEDUC2 == "7" ] <- NA
nhanes_subset$DMDEDUC2[nhanes_subset$DMDEDUC2 == "9" ] <- NA

nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2 == "7" ] <- NA
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2 == "9" ] <- NA

#delete NAs
nhanes_subset <- na.omit(nhanes_subset)
nhanes_subset_table <- na.omit(nhanes_subset)

#filter age
nhanes_subset_table <- nhanes_subset_table %>%
  filter(RIDAGEYR >= 21 & RIDAGEYR < 65)

nhanes_subset <- nhanes_subset %>%
  filter(RIDAGEYR >= 21 & RIDAGEYR < 65)

table(nhanes_subset$DMDEDUC2, nhanes_subset$RIAGENDR)
```

```{r}
#change to factor
nhanes_subset$RIAGENDR <- as.factor(nhanes_subset$RIAGENDR)
nhanes_subset$RIDRETH3 <- as.factor(nhanes_subset$RIDRETH3)
```


$~$

# Part 1. Correlation or Linear Regression
$~$

## Introduction


$~$

## Methods 


The raw data for 
$~$


## Result
Table 1 shows the first ten rows of the NHANES  2017-2018 dataset
```{r}
#Rename the categorical variable
nhanes_subset_table$RIAGENDR[nhanes_subset_table$RIAGENDR==1] <- "Male"
nhanes_subset_table$RIAGENDR[nhanes_subset_table$RIAGENDR==2] <- "Female"

nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==1] <- "Mexican American"
nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==2] <- "Other Hispanic"
nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==3] <- "Non-Hispanic White"
nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==4] <- "Non-Hispanic Black"
nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==6] <- "Non-Hispanic Asian"
nhanes_subset_table$RIDRETH3[nhanes_subset_table$RIDRETH3==7] <- "Other Race - Including Multi-Racial"

nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==1] <- "Less than 9th grade"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==2] <- "9-11th grade (incl. 12th grade with no diploma)"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==3] <- "High school graduate/GED or equivalent"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==4] <- "Some college or AA degree"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==5] <- "College graduate or above"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==7] <- "Refused"
nhanes_subset_table$DMDEDUC2[nhanes_subset_table$DMDEDUC2==9] <- "Don't Know"

head(nhanes_subset_table, 10) %>%
  flextable(cwidth = 2) %>%
  fit_to_width(max_width = 7) %>%
  set_header_labels(SEQN = "Participant ID", RIAGENDR = "Gender", 
                    RIDAGEYR = "Age", BMXWT = "Weight (kg)", BMXHT 
                    ="Height (cm)", BMXBMI = "BMI (kg/m\u00B2)", SLD012 = "Sleep duration
                    (hours)", LBXHSCRP = "hs-CRP (mg/L)", 
                    LBDHDD = "HDL-Cholesterol (mg/dL)", RIDRETH3 = "Race", DMDEDUC2 = "Education level") %>%
  set_caption("First ten rows of the NHANES 2017-2018 data") %>%
  colformat_num(j = "SEQN", big.mark = "") %>%
  align_nottext_col(align = 'center') %>%
  fontsize(part = "all", size = 8) %>%
  align_text_col(align = 'center')
```

```{r}
#Calculate Mean±SD and Median
mean_bmi_crp <- rbind(mean(nhanes_subset$BMXBMI), 
                        mean(nhanes_subset$LBXHSCRP))
median_bmi_crp <- rbind(median(nhanes_subset$BMXBMI), 
                          median(nhanes_subset$LBXHSCRP))
sd_bmi_crp <- rbind(sd(nhanes_subset$BMXBMI), 
                      sd(nhanes_subset$LBXHSCRP))


#combine results and create a table
centraltendency_bmi_crp <- cbind(mean_bmi_crp, median_bmi_crp, sd_bmi_crp)
colnames(centraltendency_bmi_crp) <- c('Mean', 'Median', 'SD')
rownames(centraltendency_bmi_crp) <- c('BMI (kg/m\u00B2)','hs-CRP (mg/L)')
```

The measures of central tendency of BMI and sleep duration
```{r}
#table for measures of central tendency (table 2)
centraltendency_bmi_crp%>%
  as.data.frame %>%
  round(digits = 2) %>%
  rownames_to_column(var = " ") %>%
  flextable(cwidth = 2.5) %>%
  fit_to_width(max_width = 7) %>%
  autofit() %>%
  set_caption("The measures of central tendency of 
              Body Mass Index (BMI) and high sensitivity C-reactive protein  (hs-CRP)")
```

Several methods were used to test the normality of the data
a. Histogram
```{r}
#Histogram BMI
ggplot(nhanes_subset, aes(BMXBMI)) +
  geom_histogram(binwidth = 1, fill = "#386cb0") +
  labs(x = "BMI", caption = "Figure 1. Histogram for BMI") +
  theme(plot.caption = element_text(hjust = 0.5))

#Histogram hs-CRP
ggplot(nhanes_subset, aes(LBXHSCRP)) +
  geom_histogram(binwidth = 2, fill = "#386cb0") +
  labs(x = "hs-CRP", caption = "Figure 2. Histogram for hs-CRP" ) +
  theme(plot.caption = element_text(hjust = 0.5))
```

b. Shapiro-Wilk test
```{r}
shapiro.test(nhanes_subset$BMXBMI)
shapiro.test(nhanes_subset$LBXHSCRP)

#Shapiro-Wilk test for BMI
shapiro.test(nhanes_subset$BMXBMI) %>%
  as_flextable() %>%
  fit_to_width(max_width = 7) %>%
  colformat_double(j = 2, big.mark = "", digits = 4) %>%
  colformat_double(j = 1, big.mark = "", digits = 2) %>%
  set_caption("Shapiro-Wilk test result for BMI")

#Shapiro-Wilk test for hs-CRP
shapiro.test(nhanes_subset$LBXHSCRP) %>%
  as_flextable() %>%
  fit_to_width(max_width = 7) %>%
  colformat_double(j = 2, big.mark = "", digits = 4) %>%
  colformat_double(j = 1, big.mark = "", digits = 2) %>%
  set_caption("Shapiro-Wilk test result for hs-CRP")
```

c. Q-Q plot
```{r}
#QQplot BMI
ggplot(nhanes_subset, aes(sample = BMXBMI)) +
  stat_qq() +
  stat_qq_line() +
  labs(caption = "Figure 3. Q-Q plot of BMI") + 
  theme(plot.caption = element_text(hjust = 0.5))

#QQplot hs-CRP
ggplot(nhanes_subset, aes(sample = LBXHSCRP)) +
  stat_qq() +
  stat_qq_line() +
  labs(caption = "Figure 4. Q-Q plot of hs-CRP") +
  theme(plot.caption = element_text(hjust = 0.5))
```

We could not assume normality of the data based on the result of the histograms and Shapiro-Wilk tests. Thus, to assess the correlation between BMI and hs-CRP, Spearman’s Rank-Order correlation test was used.
$~$

Hypothesis
$~$

The null hypothesis: There is no significant correlation between BMI and hs-CRP
$~$

The alternative hypothesis: There is a significant correlation between BMI and hs-CRP
$~$

Spearman’s rank correlation test
```{r}
spearman_bmi_crp <- cor.test(nhanes_subset$BMXBMI, nhanes_subset$LBXHSCRP,
method = "spearman", exact = FALSE)

spearman_bmi_crp
```

```{r}
#table output of spearman result
spearman_bmi_crp %>%
  as_flextable() %>%
  autofit() %>%
  colformat_double(j = 1, big.mark = "", digits = 2) %>%
  colformat_double(j = 2, big.mark = "", digits = 2) %>%
  colformat_double(j = 3, big.mark = "", digits = 2) %>%
  set_caption("Result of Spearman's rank correlation test between BMI and hs-CRP")
```
Since the result of Spearman’s rank correlation test showed a p-value of less than 0.05 (p-value < 2.2e-16), the null hypothesis is rejected. Hence, we found a significant correlation between BMI and hs-CRP in adults aged 21 to 65; The Spearman’s rank correlation coefficient was 0.53, indicating a strong positive relationship.


Scatterplot to show the relationship between BMI and hs-CRP
```{r}
attach(nhanes_subset)
plot(BMXWAIST, LBXHSCRP, main = 'Scatterplot of ', xlab = 'Waist', ylab = 'CRP')
abline(lm(LBXHSCRP ~ BMXWAIST),col='red',lwd=2)
detach(nhanes_subset)

ggplot(nhanes_subset, aes(x = BMXBMI, y = LBXHSCRP)) +
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "BMI (kg/m\u00B2)", y = "hs-CRP (mg/L)",
  caption = "Figure 5. Scatterplot of BMI and hs-CRP in adults aged 21 - 65 years old") +
  stat_cor(method = "spearman") +
  theme(plot.caption = element_text(hjust = 0.5))
```

$~$


## Discussion


## References
$~$

$~$

-------------------------------------------------------------------------

# Part 2. t-test or non-parametric alternatives
$~$


## Introduction
$~$

## Methods
$~$

## Result
Research question: Is there a difference in sleeping duration between male and female?
```{r}
#check data sleep duration
summary(nhanes_subset$SLD012)
ggplot(nhanes_subset, aes(SLD012)) +
  stat_boxplot(geom ="errorbar", width = 0.5) +
  geom_boxplot()

#summary of the medians and interquartile range by group
tapply(nhanes_subset$SLD012, nhanes_subset$RIAGENDR, median)
```

Checking the assumptions
1. Test the normality of the data
a. Histograms
```{r}
nhanes_subset$RIAGENDR <- as.factor(nhanes_subset$RIAGENDR)

gender.labs <- c("Male", "Female")
names(gender.labs) <- c("1", "2")

#Histograms for sleeping duration of male and female participants
ggplot(nhanes_subset, aes(SLD012, fill = RIAGENDR)) +
  geom_histogram(bins = 25) +
  labs(x = "Sleeping duration", 
       caption = "Figure 6. Histograms for sleeping duration of male and female participants in NHANES 2017-2018") +
  theme(plot.caption = element_text(hjust = 0.5)) +
  facet_wrap(RIAGENDR ~ ., labeller = labeller(RIAGENDR = gender.labs)) +
  scale_fill_manual(values = c("#386cb0", "#ffff99"), guide="none")
```

b. Shapiro-Wilk tests
```{r}
shapiro.test(nhanes_subset$SLD012[nhanes_subset$RIAGENDR == "1"])
shapiro.test(nhanes_subset$SLD012[nhanes_subset$RIAGENDR == "2"])
```

```{r}
#table output for shapiro-wilk tests ----Male------
shapiro.test(nhanes_subset$SLD012[nhanes_subset$RIAGENDR == "1"]) %>%
  as_flextable() %>%
  fit_to_width(max_width = 7) %>%
  colformat_double(j = 1, big.mark = "", digits = 3) %>%
  colformat_double(j = 2, big.mark = "", digits = 2) %>%
  set_caption("Shapiro-Wilk Test Result for Male's Sleeping Duration")
```

```{r}
#table output for shapiro-wilk tests ----Female------
shapiro.test(nhanes_subset$SLD012[nhanes_subset$RIAGENDR == "2"]) %>%
  as_flextable() %>%
  fit_to_width(max_width = 7) %>%
  colformat_double(j = 1, big.mark = "", digits = 3) %>%
  colformat_double(j = 2, big.mark = "", digits = 2) %>%
  set_caption("Shapiro-Wilk Test Result for Female's Sleeping Duration")
```

The histograms of the sleeping duration of male and female participants display that the data is not normally distributed. In addition, the result of Shapiro-Wilk tests shows that both of the data is skewed (p-value < 2.2e-16 for both male and female data).

2. Check the homogeneity of variance
Levene's test
```{r echo=TRUE, warning=FALSE}
leveneTest(SLD012 ~ RIAGENDR, data = nhanes_subset, location = 'mean')
```

```{r echo=FALSE, warning=FALSE}
leveneTest(SLD012 ~ RIAGENDR, data = nhanes_subset, location = 'mean') %>%
  rownames_to_column(var = " ") %>%
  flextable() %>%
  colformat_double(digits = 2) %>%
  fit_to_width(max_width = 7) %>%
  set_caption("Levene's Test for Homogeneity of Variance") %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  set_header_labels(Df = "df")
```
The Levene's test p-value = 0.85 which is greater than 0.05; thus, equality of variances for the two independent samples can be assumed.

Since the normality could not be assumed for this data set, a non-parametric test (Wilcoxon Rank-Sum test / Mann-Whitney U test) was used.
$~$


Wilcoxon Rank-Sum test / Mann-Whitney U test
$~$

Hypothesis
$~$
Q: Is there a difference in sleeping duration between male and female?
The null hypothesis: there is no significant difference in sleeping duration between male and female participants in NHANES 2017-2018
$~$

The alternative hypothesis: there is a significant difference in sleeping duration between male and female participants in NHANES 2017-2018
$~$

```{r}
wilcox.test(SLD012 ~ RIAGENDR, data = nhanes_subset)
```

```{r}
wilcox_ranksum_gender_sleep <- wilcox.test(SLD012 ~ RIAGENDR, data = nhanes_subset)

wilcox_ranksum_gender_sleep %>%
  as_flextable() %>%
  colformat_double(j = 1, big.mark = "", digits = 0) %>%
  colformat_double(j = 2, big.mark = "", digits = 2) %>%
  set_caption("Wilcoxon Rank-Sum test result for difference of sleeping duration between male and female in NHANES 2017-2018")

```

The result of Wilcoxon’s rank-sum test / Mann-Whitney U test showed that W = 1268260 with p-value = 2.831e-08. Thus, there was a significant difference between the sleeping duration of male and female aged 21-64 years old in this study.

The measures of central tendency of sleeping duration based on the gender
```{r}
nhanes_subset_table$SLD012 <- as.numeric(nhanes_subset_table$SLD012)

centraltendency_sleep_gender <- group_by(nhanes_subset_table, RIAGENDR) %>%
  summarise(
    count = n(),
    mean = mean(SLD012, na.rm = TRUE),
    median = median(SLD012, na.rm = TRUE),
    sd = sd(SLD012, na.rm = TRUE)
    )

#change the gender label from 1 & 2 to Male and Female in the output table
centraltendency_sleep_gender$RIAGENDR[centraltendency_sleep_gender$RIAGENDR==1] <- "Male"
centraltendency_sleep_gender$RIAGENDR[centraltendency_sleep_gender$RIAGENDR==2] <- "Female"
  

#output table for measures of central tendency
centraltendency_sleep_gender %>%
  flextable %>%
  set_header_labels(RIAGENDR = "Gender", count = "Count", mean = "Mean", median = "Median",
                    sd = "SD") %>%
  colformat_double(j = 3, big.mark = "", digits = 2) %>%
  colformat_double(j = 4, big.mark = "", digits = 2) %>%
  colformat_double(j = 5, big.mark = "", digits = 2) %>%
  set_caption("The measures of central tendency of the sleeping duration of male and female participants in NHANES 2017-2018")
```
Boxplots
```{r}
#Boxplots
ggplot(nhanes_subset_table, aes(x = RIAGENDR, y = SLD012, group = RIAGENDR)) +
  xlim("1" = "Male", "2" = "Female") +
  stat_boxplot(geom ="errorbar", width = 0.5) +
  geom_boxplot() +
  labs(x = "Gender", y = "Sleeping duration (hours)", caption = "Figure 7. Boxplots of the sleeping duration of male and female participants in NHANES 2017-2018") +
  theme(plot.caption = element_text(hjust = 0.5))
```
The boxplots show that the median sleeping duration of female participants is slightly higher than male participants. In comparison to the male boxplot, the female boxplot is relatively tall; it suggests that female participants have more varied sleeping durations. There are some outliers in both male and female sleeping duration data.
$~$
 
## Discussion
The Wilcoxon’s rank-sum test showed a significant difference (W = 1268260, p-value = 2.831e-08) of sleeping duration between male and female participants of NHANES 2017-2018. The median sleeping duration of male participants was 7.00 hours compared to 7.50 hours for female participants.


## References

$~$

$~$

-------------------------------------------------------------------------

# Part 3. ANOVAs or non-parametric alternatives
$~$

## Introduction

$~$

## Methods
The Salaries data fr

$~$

## Result
The measures of central tendency of HDL-Cholesterol level based on the race
```{r}
nhanes_subset_table$LBDHDD <- as.numeric(nhanes_subset_table$LBDHDD)

centraltendency_race_hdl <- group_by(nhanes_subset_table, RIDRETH3) %>%
  summarise(
    count = n(),
    mean = mean(LBDHDD, na.rm = TRUE),
    median = median(LBDHDD, na.rm = TRUE),
    sd = sd(LBDHDD, na.rm = TRUE)
  )
  
#output table for measures of central tendency
centraltendency_race_hdl %>% 
  flextable %>%
  set_header_labels(RIDRETH3 = "Race", count = "Count", mean = "Mean", median = "Median", sd = "SD") %>%
  autofit() %>%
  colformat_double(j = 3, big.mark = "", digits = 2) %>%
  colformat_double(j = 4, big.mark = "", digits = 2) %>%
  colformat_double(j = 5, big.mark = "", digits = 2) %>%
  set_caption("The measures of central tendency of HDL-Cholesterol level based on the race")
```



Boxplot --------- DO I NEED THIS? ---------------
```{r}
#Boxplots
ggplot(nhanes_subset, aes(x = RIDRETH3, y = LBDHDD, fill = RIDRETH3)) +
  stat_boxplot(geom ="errorbar", width = 0.5) + 
  geom_boxplot() +
  labs(x = "Race", y = "HDL-Cholesterol (mg/dL)", caption = "Figure XX. Boxplots of HDL-Cholesterol level among several different races of NHANES 2017-2018 participants aged 21 - 65 years old") +
  theme(plot.caption = element_text(hjust = 0.1),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
   scale_fill_manual(values = c("#f0f9e8", "#ccebc5", "#a8ddb5", "#7bccc4", "#43a2ca",
                               "#0868ac"), name = "RIDRETH3", 
                     labels = c("Mexican American", "Other Hispanic", "Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Asian", "Other Race - Including Multi-Racial")) +
  guides(fill = guide_legend(title= "Race"))
```

ANOVA

Hypothesis
$~$

The null hypothesis: there is no significant difference of HDL-Cholesterol level among difference races of people aged 21 - 65 years old in NHANES 2017-2018
$~$

The alternative hypothesis: there is a significant difference of HDL-Cholesterol level among difference races of people aged 21 - 65 years old in NHANES 2017-2018
$~$

```{r}
nhanes_subset$LBDHDD <- as.numeric(nhanes_subset$LBDHDD)

anova_hdl_race <- aov(LBDHDD ~ RIDRETH3, data = nhanes_subset)
summary(anova_hdl_race)
```

Table XX shows the result of the one-way ANOVA between HDL-Cholesterol level and race
```{r}
#output table for anova
anova_hdlrace_df <- c(5, 3485)
anova_hdlrace_SumSq <- c(22848, 856520)
anova_hdlrace_MeanSq <- c(4570, 246)
anova_hdlrace_Fvalue <- c(18.59, "")
anova_hdlrace_pvalue <- c("<2e-16", "")

anova_hdlrace_table <- cbind(anova_hdlrace_df,
                             anova_hdlrace_SumSq,
                             anova_hdlrace_MeanSq, 
                             anova_hdlrace_Fvalue, 
                             anova_hdlrace_pvalue)

colnames(anova_hdlrace_table) <- c('df', 'Sum Sq', 'Mean Sq', 'F value', 'P value')
rownames(anova_hdlrace_table) <- c('Race','Residuals')

anova_hdlrace_table %>%
  as.data.frame() %>%
  rownames_to_column(var = " ") %>%
  flextable()  %>%
  autofit() %>%
  fit_to_width(max_width = 7) %>%
  set_caption("Result of One-way ANOVA")
```


It is necessary to see whether the data fit the assumptions for one-way ANOVA:
$~$

1. The residuals should be normally distributed
```{r}
#Extract the residuals
anova_hdlrace_res_compare <- residuals(object = anova_hdl_race)
anova_hdlrace_res_compare

anova_hdlrace_residuals <- as_data_frame(anova_hdl_race$residuals)
anova_hdlrace_residuals
```

```{r}
#histogram
ggplot(anova_hdlrace_residuals, aes(value)) +
  geom_histogram(bins = 50) +
  labs(x = "Standardised residuals", 
       caption = "Figure XX. Histogram of the standardised residuals") +
   theme(plot.caption = element_text(hjust = 0.5))
```

```{r echo=FALSE}
#QQ plot
plot(anova_hdl_race, which = 2, caption = " ", sub = "Figure XX. Q-Q plot of standardised residuals")
```

```{r echo=TRUE}
#Shapiro-Wilk test of normality
shapiro.test(x = anova_hdl_race$residuals)
```

```{r echo=FALSE, warning=FALSE}
#Shapiro-Wilk table
shapiro.test(x = anova_hdl_race$residuals) %>%
  as_flextable() %>%
  colformat_double(digits = 2) %>%
  fit_to_width(max_width = 7) %>%
  set_caption("Shapiro-Wilk test result of the residuals") %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center')
```

The histogram and QQ plot displays that the residuals skewed to the right. In addition, the Shapiro-Wilk test shows that the p-value is significant (p-value < 2.2e-16). Hence, we cannot assume normality of the residuals.

2. Homogeneity of variance
$~$

The variances should be similar for all groups
$~$

```{r echo=FALSE}
plot(anova_hdl_race, 1, sub.caption = "Figure XX. Residuals vs Fitted values plot", caption = " ")
```

The result of Residuals vs Fitted plot shows that the spread of the residuals in some of the group are almost more than three times the spread in the group with the smallest spread.

```{r eval=FALSE, include=FALSE}
#function of residuals vs fitted values
#This allows you to see if the variability of the observations differs across the groups #because all observations in the same group get the same fitted value.
```

Levene test
```{r echo=TRUE, warning=FALSE}
leveneTest(LBDHDD ~ RIDRETH3, data = nhanes_subset)
```

```{r echo=FALSE, warning=FALSE}
leveneTest(LBDHDD ~ RIDRETH3, data = nhanes_subset) %>%
  rownames_to_column(var = " ") %>%
  flextable() %>%
  colformat_double(digits = 2) %>%
  fit_to_width(max_width = 7) %>%
  set_caption("Levene's Test for Homogeneity of Variance (center = median)") %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  set_header_labels(Df = "df")
```

The result of Levene's test shows a significant result (p = 1.327e-05); there is evidence to suggest that the variance across groups is statistically significantly different. Therefore, we cannot assume the homogeneity of variances in the six different race groups.

Since the ANOVA assumptions were not met, a non-parametric alternative to one-way ANOVA;Kruskal-Wallis test was used.

```{r echo=TRUE}
#Kruskal-Wallis test
kruskal.test(LBDHDD ~ RIDRETH3, data = nhanes_subset)
```
Table XX shows the result of Kruskal-Wallis rank sum test between HDL-Cholesterol level and race
```{r}
#Output table for Kruskal-Wallis
kruskal.test(LBDHDD ~ RIDRETH3, data = nhanes_subset) %>%
  as_flextable() %>%
  colformat_double(digits = 2) %>%
  autofit() %>%
  fit_to_width(max_width = 7) %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  set_caption("Result of Kruskal-Wallis rank sum test")
```
As the p-value < 0.001, there was strong evidence to suggest a difference between at least one pair of the groups. In order to discover which pairs, we conducted the pairwise Wilcoxon signed-rank comparisons for each pair of groups.

Pairwise Comparisons
Since multiple comparisons were being done, the Bonferroni adjustment was used to keep the overall type 1 error low.
```{r echo=TRUE, warning=FALSE}
#Pairwise Comparisons
pairwise.wilcox.test(nhanes_subset$LBDHDD, nhanes_subset$RIDRETH3, p.adj = "bonferroni", exact = F)
```

```{r}
#Output table for Pairwise Wilcoxon
PairwiseWilcox_1 <- c(1.0000, 0.0157, 1.1e-15, 0.0017, 1.0000)
PairwiseWilcox_2 <- c("-", 1.0000, 3.6e-08, 0.2554, 1.0000)
PairwiseWilcox_3 <- c("-", "-", 1.2e-07, 1.0000, 1.0000)
PairwiseWilcox_4 <- c("-", "-", "-", 0.0021, 2.9e-05)
PairwiseWilcox_6 <- c("-", "-", "-", "-", 0.7297)

PairwiseWilcox_table <- cbind(PairwiseWilcox_1,
                             PairwiseWilcox_2,
                             PairwiseWilcox_3, 
                             PairwiseWilcox_4, 
                             PairwiseWilcox_6)

colnames(PairwiseWilcox_table) <- c('Mexican American', 'Other Hispanic', 'Non-Hispanic 
                                    White', 'Non-Hispanic Black', 'Non-Hispanic Asian')
rownames(PairwiseWilcox_table) <- c('Other Hispanic', 'Non-Hispanic White', 'Non-Hispanic 
                                    Black', 'Non-Hispanic Asian', 'Other Race - Including 
                                    Multi-Racial')


PairwiseWilcox_table %>%
  as.data.frame() %>%
  rownames_to_column(var = " ") %>%
  flextable()  %>%
  fit_to_width(max_width = 7) %>%
  autofit() %>%
  fontsize(part = "all", size = 10) %>%
  set_caption("Pairwise comparisons using Wilcoxon rank sum test with continuity correction") %>%
  footnote(i = 1, value = as_paragraph(c("P value adjustment method: bonferroni ")), ref_symbols = c(""))
```
Table XX shows the results of the Wilcoxon tests on each pair of groups. The adjusted p-values are shown for each pair in the output. There were significant differences between:
(1) Non-Hispanic White and Mexican American (p-value = 0.0157), (2) Non-Hispanic Black and Mexican American (p-value = 1.1e-15), (3) Non-Hispanic Asian and Mexican American (p-value = 0.0017), (4) Non-Hispanic Black and Other Hispanic (p-value = 3.6e-08), (5) Non-Hispanic Black and Non-Hispanic White (p-value = 1.2e-07), (6) Non-Hispanic Asian and Non-Hispanic Black (p-value = 0.0021), (7) Other Race - Including Multi-Racial and Non-Hispanic Black (p-value = 2.9e-05).

```{r echo=FALSE, warning=FALSE}
#Visualisation Pairwise comparisons using Wilcoxon rank sum test
PairwiseWilcox_comparisons_ggboxplot <- list(c("1", "3"), c("1", "4"), c("1", "6"), c("2", "4"), c("3", "4"), c("4", "6"), c("4", "7"))

vis_pairwise <- 
  ggboxplot(nhanes_subset, x = "RIDRETH3", y = "LBDHDD", color = "RIDRETH3", 
            bxp.errorbar = TRUE) + 
  stat_compare_means(comparisons = PairwiseWilcox_comparisons_ggboxplot, 
                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), 
                                        symbols = c("****", "***", "**", "*", "ns")),) +
  stat_compare_means(label.y = 325) + # Add global Kruskal_Wallis p-value
  scale_color_discrete(labels = c('1: Mexican American', '2: Other Hispanic', '3: Non-Hispanic White','4: Non-Hispanic Black', '6: Non-Hispanic Asian', 
                                  '7: Other Race - Including Multi-Racial'))

#set title and y label
ggpar(vis_pairwise,  xlab = "Race", ylab = "HDL-Cholesterol (mg/dL)", 
      legend = "bottom", legend.title = "Race", ticks = FALSE, font.x = 11, font.y = 11, 
      font.tickslab = 9, caption = "Figure XX. Visualisation of the significant result of Pairwise comparisons
      **: p ≤ 0.01, ***: p ≤ 0.001, ****: p ≤ 0.0001")

```
## Discussion
A Kruskal-Wallis test was carried out to compare the HDL-Cholesterol level among six races groups of the NHANES participants. There was robust evidence of difference (p-value < 2.2e-16) among the mean ranks of at least one pair of groups. Wilcoxon signed-rank pairwise tests were conducted for the six pairs of groups. There were strong evidence (p-value < 0.05, adjusted using the Bonferroni correction) of significant differences in seven comparisons among the groups.

The median HDL-Cholesterol level for the Non-Hispanic Black was 54 mg/dL compared to 50 mg/dL for Non-Hispanic Asian. Further, Non-Hispanic White and Other Race - including Multi-Racial have the same median HDL-Cholesterol level of 49 mg/dL. Lastly, Mexican American and Other Hispanic have the lowest median HDL-Cholesterol level of 48 mg/dL.


$~$

## References

$~$

-------------------------------------------------------------------------

# Part 4. Chi-square test
$~$



## Introduction

Research question: Is there any association between BMI and education level among NHANES 2017-2018 participants?
$~$

## Methods

The education and BMI data from NHANES 2017-2018 were used for this study. The data on education consisted of five categories - Less than 9th grade, 9-11th grade (including 12th grade with no diploma), High school graduate/GED or equivalent, Some college or AA degree, and College graduate or above. The BMI data (continuous data) was transformed into a categorical variable based on the Centers for Disease Control and Prevention (CDC) definition. Thus, the BMI was categorised into four groups by using as.factor() and ifelse() functions.

The contingency table with observed frequencies was obtained using the table() command, and the column percentages were derived by using the prop.table() command. Pearson’s Chi-squared test was used to analyse whether there is any significant association between BMI and education level among the participants of NHANES 2017-2018. Additionally, a bar plot was made to give a graphical visualisation.

$~$

## Result

```{r}
#change chr to factor
nhanes_subset$DMDEDUC2 <- as.factor(nhanes_subset$DMDEDUC2)

#continuous BMI to categorical BMI
nhanes_subset$bmi_categorical <- as.factor(
  ifelse(nhanes_subset$BMXBMI < 18.5, '1',
  ifelse(nhanes_subset$BMXBMI >= 18.5 & nhanes_subset$BMXBMI < 25, '2',
  ifelse(nhanes_subset$BMXBMI >= 25 & nhanes_subset$BMXBMI < 30, '3',
  ifelse(nhanes_subset$BMXBMI >= 30, '4', 0)))))

#rename the BMI categorical levels
levels(nhanes_subset$bmi_categorical) <- c("Underweight", "Healthy weight", "Overweight", "Obese")

#rename the education categorical levels
levels(nhanes_subset$DMDEDUC2) <- c("Less than 9th grade", "9-11th grade (incl. 12th grade with no diploma)", "High school graduate/GED or equivalent", "Some college or AA degree", "College graduate or above")
```

Contingency table containing the observed frequencies
```{r echo=TRUE, warning=FALSE}
# BMI & education level -- contingency table
attach(nhanes_subset)
ct_bmi_educ <- table(bmi_categorical, DMDEDUC2)
addmargins(ct_bmi_educ)
detach(nhanes_subset)
```

```{r}
#output table of contingency table -----OBSERVED FREQUENCIES----------
bmi_educ_1 <- c(2, 36, 109, 111, 258)
bmi_educ_2 <- c(8, 93, 107, 169, 377)
bmi_educ_3 <- c(15, 200, 228, 367, 810)
bmi_educ_4 <- c(13, 239, 332, 586, 1170)
bmi_educ_5 <- c(13, 271, 303, 289, 876)
bmi_educ_sumcol <- c(51, 839, 1079, 1522, 3491)

bmi_educ_table <- cbind(bmi_educ_1, bmi_educ_2, bmi_educ_3, bmi_educ_4,
                        bmi_educ_5, bmi_educ_sumcol)

colnames(bmi_educ_table) <- c('Less than 9th grade', 
                              '9-11th grade (incl. 12th grade with no diploma)',
                              'High school graduate/GED or equivalent', 
                              'Some college or AA degree',
                              'College graduate or above', 'Sum')

rownames(bmi_educ_table) <- c('Underweight', 'Healthy weight', 'Overweight', 'Obese', 
                              'Sum')

bmi_educ_table %>%
  as.data.frame() %>%
  rownames_to_column(var = " ") %>%
  flextable()  %>%
  fit_to_width(max_width = 7) %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  fontsize(size = 9, part = "all") %>%
  autofit() %>%
  set_caption("Contingency table containing the observed frequencies")
```
Table XX shows the contingency table of observed frequencies. Almost half of the participants were obese (1,522 people). The subjects with some college or AA degree had the highest number of overweight (332) and obese (586) people. Most of the college graduates or above were overweight (303 people); However, this group also had the highest number of people with healthy weights compared to the other groups with lower education levels.
$~$

```{r}
percentage_bmi_educ <- round(100*prop.table(ct_bmi_educ, 2), digits = 2)
percentage_bmi_educ
```

```{r}
#Result table
#Contingency table -----PERCENTAGES----------
percentage_bmi_educ_1 <- c(0.78 , 13.95, 42.25, 43.02)
percentage_bmi_educ_2 <- c(2.12, 24.67, 28.38, 44.83)
percentage_bmi_educ_3 <- c(1.85, 24.69, 28.15, 45.31)
percentage_bmi_educ_4 <- c(1.11, 20.43, 28.38, 50.09)
percentage_bmi_educ_5 <- c(1.48, 30.94, 34.59, 32.99)

percentage_bmi_educ_table <- cbind(percentage_bmi_educ_1,
                             percentage_bmi_educ_2,
                             percentage_bmi_educ_3,
                             percentage_bmi_educ_4,
                             percentage_bmi_educ_5)

colnames(percentage_bmi_educ_table) <- c('Less than 9th grade', 
                                         '9-11th grade (incl. 12th grade with no diploma)', 
                                         'High school graduate/GED or equivalent', 
                                         'Some college or AA degree',
                                         'College graduate or above')

rownames(percentage_bmi_educ_table) <- c('Underweight', 'Healthy weight', 'Overweight', 'Obese')


percentage_bmi_educ_table %>%
  as.data.frame() %>%
  rownames_to_column(var = " ") %>%
  flextable()  %>%
  autofit() %>%
  fit_to_width(max_width = 7) %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  fontsize(size = 10) %>%
  fontsize(size = 10, part = "header") %>%
  set_caption("Contingency table of percentages")
```

The contingency table of percentages shows that more college graduates or above have the highest percentage of healthy weight (30.94%). People with less than 9th-grade education level are 42.25% overweight, and 50.09% of participants with some college or AA degree are obese.

Hypothesis
$~$

The null hypothesis: Participants' education level is not associated with BMI
$~$

The alternative hypothesis: Participants' education level is associated with BMI
$~$


```{r}
attach(nhanes_subset)
chisq_bmi_educ <- chisq.test(table(DMDEDUC2, bmi_categorical))
chisq_bmi_educ
detach(nhanes_subset)
```

```{r}
#Chi-squared test result table
chisq_bmi_educ %>%
  as_flextable() %>%
  fit_to_width(max_width = 7) %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  colformat_double(j = 2, big.mark = "", digits = 2) %>%
  colformat_double(j = 1, big.mark = "", digits = 2) %>%
  set_caption("Result of the Pearson's Chi-squared test")
```

Pearson's Chi-squared test results are shown in Table XX. The null hypothesis must be rejected because there was a significant outcome ($\chi 2$ = 92.97, df = 12, p = 1.308e-14). According to this finding, BMI was significantly associated with education level of the participants.

```{r echo=FALSE, warning=TRUE}
#observed freq as a dataframe used for bar graph
attach(nhanes_subset)
bmi_educ_vis <- table(DMDEDUC2, bmi_categorical)
bmi_educ_vis <- as.data.frame(bmi_educ_vis)
bmi_educ_vis
detach(nhanes_subset)
```

```{r echo=FALSE, warning=TRUE}
#Clustered bar chart 1 -----------USE THIS-----------
ggplot(bmi_educ_vis, aes(x = DMDEDUC2, y = Freq, fill = bmi_categorical)) +
  geom_bar(stat="identity", position = position_dodge()) +
  labs(x = "Education level", y = "Frequency", fill = "BMI", 
       caption = "Figure XX. The clustered bar chart shows the frequency of BMI of the NHANES 2017-2018 
participants according to their education level") +
  theme(plot.caption = element_text(hjust = 0.0), 
        axis.text.x = element_text(size = 8)) +
  scale_x_discrete(labels = c("Less than \n 9th grade", "9-11th \n grade (incl. 12th grade \n with no diploma)",  "High school \n graduate/GED \n or equivalent", "Some college \n or AA degree", "College graduate \n or above")) +
  scale_fill_manual(values = c("Underweight" = "#ffffb3", "Healthy weight" = "#ccebc5", "Overweight" = "#bebada", "Obese" = "#fb8072"))
```


```{r}
#Clustered bar chart 2
ggplot(bmi_educ_vis, aes(x = bmi_categorical, y = Freq, fill = DMDEDUC2)) +
  geom_bar(stat="identity", position = position_dodge()) +
  labs(x = "BMI", y = "Frequency", fill = "Education level", 
       caption = "Figure XX. The clustered bar chart shows the frequency of BMI of the 
       NHANES 2017-2018 participants according to their education level. Most of --") +
  theme(plot.caption=element_text(hjust = 0.0)) +
   scale_fill_manual(values = c("Less than 9th grade" = "#fb8072", "9-11th grade (Incl. 12th grade with no diploma)" = "#fed9a6", "High school graduate/GED or equivalent" = "#ffffb3", "Some college or AA degree" = "#bebada", "College graduate or above" = "#ccebc5"))
```

To check on the validity of chi-squared test, we need to examine if one or more of the expected frequencies are less than 5.

```{r echo=TRUE, warning=FALSE}
chisq_bmi_educ_exp <- chisq_bmi_educ$expected
chisq_bmi_educ_exp
```

```{r Table_10, echo=FALSE, warning=FALSE}
#Result table of the expected frequencies
chisq_bmi_educ_exp %>%
  as.data.frame() %>%
  round(digits = 2) %>%
  rownames_to_column(var = " ") %>%
  flextable() %>%
  autofit() %>%
  fit_to_width(max_width = 7) %>%
  align_nottext_col(align = 'center') %>%
  align_text_col(align = 'center') %>%
  set_caption("Result of expected frequencies")
```
From the table above, there is only one cell with the expected count of less than 5. According to (Yates, Moore & McCabe, 1999, p. 734), "no more than 20% of the expected counts are less than five, and all individual expected counts are one or greater"(1). Since only one out of twenty cells had an expected count of less than 5, the Chi-squared test was valid.
$~$


## Discussion

A Pearson's Chi-squared analysis was conducted to analyse whether there was an association between participants' education level and BMI. The result showed a significant relation ($\chi 2$ = 92.97, df = 12, p = 1.308e-14) between NHANES 2017-2018 participants' education level and their BMI. Participants who were college graduates or above had the highest percentage of healthy weight (30.94%) in contrast to other groups with lower education levels. The highest overweight percentage was found in the group of people with an education level of less than 9th grade (42.25%); further, participants with some college or AA degree had the most significant percentage of obese people (50.09%). These findings align with prior studies(2,3) that education level is related to general obesity.

$~$


## References



$~$